### 2021-03-09 paper.Rmd
clustering example complete

\begin{figure*}[htp]
         \includegraphics[width=0.5\textwidth]{figures/cluster_example_r.pdf}
     \vfill
   \begin{minipage}{\linewidth}     
   \centering
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour1_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour2_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour3_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour4_r.pdf}
   \end{minipage}%
\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and projections seen during a tour (B.1-4). There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied. This can be seen in the tour projections. The two green clusters are spherical in shape, and very distance from the brown clusters. Three of the brown clusters are very close to each other (just visible in B.1), and all brown clusters are elliptical. The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the dimension reduction.}
\label{fig:tsne-tour}
\end{figure*}

>>>>

Nonlinear dimension reduction techniques such as t-SNE have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary. However, it is only a summary, and it likely involves substantial warping of the original data space. Using the tour along with these techniques can illuminate the nature of the warped space, and reveal other structure lost in the dimension reduction.  Figure \ref{fig:tsne-tour} illustrates the difference in what can be learned from a tour in comparison with nonlinear dimension reduction using t-SNE. The 10-dimensional data comes from @Rauber2009-kk. The t-SNE view (plot A) shows six clearly separated clusters, all spherical with different sizes. The tour shows that the clusters look very little like this in the full data space. The clusters are at various distances apart and very different sizes. In the four projections from the tour this can be seen. The two green clusters are large and spherical, and far from the brown clusters. The brown clusters have one larger one, and three smaller, very close to each other. All of these are elliptical, which means that they have very little variability, actually no variability if you watch the full tour, in some of the 10 dimensions. This gives some deeper perspective to what is learned from t-SNE, and illustrates what the t-SNE dimension reduction has done: it has found small gaps between points and expanded these gaps to yield the representation. It should be noted, though, that the t-SNE view clearly reveals the presence of six distinct clusters. With this information is the invitation to look closer at the data in the tour, to see, that yes, indeed, there are three, tiny, tiny clusters very close to each other.

\begin{figure*}[!t]
         \includegraphics[width=0.5\textwidth]{figures/cluster_example_r.pdf}
     \vfill
   \begin{minipage}{\linewidth}     
   \centering
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour1_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour2_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour3_r.pdf}
         \includegraphics[width=0.24\textwidth]{figures/cluster_example_tour4_r.pdf}
   \end{minipage}%
\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and projections seen during a tour (B.1-4). There are six clusters, as seen in the t-SNE view, but the relative distance between the clusters is extremely varied. This can be seen in the tour projections. The two green clusters are spherical in shape, and very distance from the brown clusters. Three of the brown clusters are very close to each other (just visible in B.1), and all brown clusters are elliptical. The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the dimension reduction.}
\label{fig:tsne-tour}
\end{figure*}
